DNNNNR0: How to Avoid OOTA Without Really Trying

### 2025-05-11 paper/ootavacuous.tex
ootavacuous: Wordsmith Sections 6.4 and 6.4.1
Signed-off-by: Paul E. McKenney <paulmck@kernel.org>

This section exercises the definition of ``semantic dependency''
presented in the previous section on selected examples.
These exercises use the following steps:

@@

Section~\ref{sec:Semantic Dependencies and Function Calls}
provides the following example:
\begin{quote}
\begin{verbatim}
y = f(x);
\end{verbatim}
\end{quote}
If the compiler knows nothing about the function \co{f()} and the variable
\co{y}, it must emit code that loads from \co{x}, passes this value to
\co{f()}, and finally stores the return value to \co{y}.
Is there a semantic dependency from the load of \co{x} to the store of
\co{y}?
To determine this, it is necessary to apply this same analysis to the
definition of \co{f()}, treating its argument as a load and each of its
\co{return} statements as a store.
If and only if this analysis shows that, for a given execution, there
is a semantic dependency within \co{f()}, then there must also be a
semantic dependency from the load of \co{x} to the store of \co{y}.
Of course, if the compiler cannot prove that there is no such semantic
dependency, it must assume than there is one.

If \co{f} is a pointer to a function, then the compiler must also
generate load from \co{f}, so that there might be a semantic dependency
from either or both of \co{f} and \co{y} to the store to \co{y}.
This is straightforward:
A given execution will load a pointer to a particular function from
\co{f}, so if there is another execution that loads a pointer to a
different function, and the two functions return different values for
that execution's value from \co{y}, then there is a semantic dependency
from \co{f} to \co{y}.

Situations where the compiler knows about \co{f} more often result in
inlining, in which case there is no function call, allowing the semantic
dependencies (or lack thereof) to be determined with respect to the
inlined code.

>>>>

This section exercises the definition of ``semantic dependency'' on
selected examples.
These exercises use the following steps:

@@

Section~\ref{sec:Semantic Dependencies and Function Calls}
provides the following example:
\begin{quote}
\begin{verbatim}
y = f(x);
\end{verbatim}
\end{quote}
If the compiler knows nothing about the function \co{f()} and the variable
\co{x}, it must emit code that loads from \co{x}, passes this value to
\co{f()}, and finally stores the return value to \co{y}.
Is there a semantic dependency from the load of \co{x} to the store of
\co{y}?
To determine this, it is necessary to apply this same analysis to the
definition of \co{f()}, treating its argument as a load and each of its
\co{return} statements as a store.
If and only if this analysis shows that, for a given execution, there
is a semantic dependency within \co{f()}, then there must also be a
semantic dependency from the load of \co{x} to the store of \co{y}.
As always, if the compiler cannot prove that there is no such semantic
dependency, it must assume than there is one.

If \co{f} is a pointer to a function, then the compiler must also emit
a load from \co{f}, so that there might also be a semantic dependency
from the load from \co{f} to the store to \co{y}.
In the context of C++, this is straightforward:
A given execution will load a pointer to a particular function from
\co{f}, so if there is another execution that loads a pointer to a
different function, and the two functions return different values for
the value loaded from \co{x}, then there is a semantic dependency from
\co{f} to \co{y}.
Languages that permit functions to be generated at runtime must use
more complicated analysis.

Situations where the compiler knows about \co{f} often result in
inlining, in which case there is no function call, allowing the semantic
dependencies (or lack thereof) to be determined with respect to the
inlined code.
